---
title: "Correlations between phenotypes"
author: "Donghyung Lee"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load packages
```{r}
rm(list=ls())
#library(data.table)
#library(reshape2)
#library(ggplot2)

library(dplyr)
library(tidyr)
library(pheatmap)
library(RColorBrewer)
```


## Read raw phenotypes
```{r}
pheno.file = "~/Box Sync/Miami_IMPC/data/v10.1/AllBehaviourControls.Rdata"
load(file=pheno.file)
pheno = df %>% filter(data_type!="TEXT") %>% 
  mutate(proc_short_name=recode(procedure_name, "Acoustic Startle and Pre-pulse Inhibition (PPI)"="PPI",
         "Combined SHIRPA and Dysmorphology"="SHIRPA",
         "Fear Conditioning"="FEAR",
         "Grip Strength"="GRIP",
         "Hole-board Exploration"="HB",
         "Light-Dark Test"="LD",
         "Open Field"="OF",
         "Rotarod"="RT",
         "Tail Suspension"="TS")) %>%
  mutate(proc_param_name=paste0(proc_short_name,"_",parameter_name))
dim(pheno)
```

## Heatmap showing measured phenotypes 

This heatmaps show phenotypes measured for each control mouse. Rows represent mice and columns represents phenotypes. 
```{r, fig.width=6, fig.height=8}
mtest <- table(pheno$biological_sample_id, pheno$proc_param_name)
mtest <-as.data.frame.matrix(mtest)
pheatmap(mtest, cluster_rows = F, cluster_cols=F, show_rownames=F, color=brewer.pal(n=9,name="Reds"))
```
SHIRPA, PPI, OF and GRIP phenotypes are measured for most of control mice. So we compute correlations of these phenotypes.

## Subsetting SHIRPA, PPI, OF, GRIP
```{r} 
pheno.sub <- pheno %>% filter(proc_short_name=="SHIRPA"|
                                proc_short_name=="PPI"|
                                proc_short_name=="OF"|
                                proc_short_name=="GRIP")
dim(pheno.sub)
```

## Reshape the data (long to wide format)
```{r}
pheno.mat <- pheno.sub %>% 
  select(biological_sample_id, proc_param_name, data_point) %>%
  spread(proc_param_name, data_point) %>%
  tibble::column_to_rownames(var="biological_sample_id")
dim(pheno.mat)
```

## Compute correlations between phenotypes
```{r, fig.width=8, fig.height=4}
pheno.cor <- cor(pheno.mat, use="pairwise.complete.obs")
pheatmap(pheno.cor, cluster_rows = F, cluster_cols=F, show_colnames=F, col=brewer.pal(n=11,name="RdBu"))
```


## ETC
```{r}
## number of domains
length(unique(pheno$procedure_name))
unique(pheno$procedure_name)
table(pheno$procedure_name)
## number of phenotypes in each domain
pheno %>% group_by(procedure_name) %>% summarise(num.pheno=length(unique(parameter_name)))
## number of observations in each domain & phenotype
pheno %>% group_by(procedure_name, parameter_name) %>% summarise(num.obs=length(parameter_name))

table(pheno$biological_model_id, pheno$strain_name)
table(pheno$zygosity)
table(pheno$sex)
table(pheno$production_center)
table(pheno$procedure_name, pheno$production_center)
```
